Charlson Comorbidity Index Calculator
=====================================

The Charlson index is a commonly used tool to assess medical commodities, and to adjust for risk in health-care services and research [1][1] [2][2]. 
Electronic medical records(EMR) are widespread today, and diagnosis codes are usually easily extracted from the EMR.
The program here processes ICD9 codes for a given number of patients, and produces a Charlson comorbidity table along with a calculated Charlson score. 

# How to use the program
The program is written in the R programming language. It requires [R][3] to be installed, along with the following packages: `reshape2`, `plyr`, and `ggplot2`.
In order to use the program, you will have to add patient identifiers, along with corresponding ICD9 codes to the `input` folder. With this current built, some test patients are added so that the program can run. Here is what you need to do:

- Replace `Patients1.csv`. This file contains patients identifiers (MRN), date of birth (DOB), and date of hospitalization (DOH). Here is how the file look like: 
```{r echo=FALSE} 
head(p1)
```

- Replace `Patients2.csv`. This file contains the ICD9 codes along with the dates when the diagnoses were made. Here is how the file looks like:
```{r echo=FALSE} 
head(p2)
```

- All the dates should be in the following format: `YYYY-MM-DD`

- The program defaults to only accepting ICD9 diagnoses that were made within 5 years of the hospitalization date. However, you can change this by changing the number in `CharlsonRules4.csv` file.

- After you change the files as outlined above, run `CharlsonCalc.R` file. Make sure that the working directory is set to where this file is. You can use `setwd` function to change the working directory. 

# Output
The information is processed, and two files are written to the `output` folder. `Charlson.csv` is a table with the patients, their comorbidities, and their Charlson score. `Graph.png` summarizes the comorbidities for all patients. Here is how the output looks like:

```{r echo=FALSE} 
Final
plot
```

Your feedback is always appreciated. Thanks!

[1]: http://www.ncbi.nlm.nih.gov/pubmed/12725876/
[2]: http://www.ncbi.nlm.nih.gov/pubmed/16015512/
[3]: http://www.r-project.org/